### 2016-06-29 text/property_verification_proofs.tex
Move property verification proofs into own section.

\section{Validity and Completeness of Property Verification}
\label{sec:val_complete_prop_verif}

\counterwithin{proposition}{section}
\counterwithin{lemma}{section}
\counterwithin{equation}{section}
\setcounter{proposition}{0}
\setcounter{lemma}{0}
\setcounter{equation}{0}

\onehalfspacing

\begin{definition}{Property of a Transformation\\}
\label{def:trans_prop_appendix}
\CatchFileBetweenTags{\propertydef}{text/definitions}{propertydef}{\propertydef}
\end{definition}

In \cref{def:trans_prop_appendix}, pre-conditions use the same pattern language as the match graph in
DSLTrans rules, allowing the possibility of including several instances of
the same metamodel element as well as indirect links in the property. Indirect links in properties have
the same meaning as in the rule match graph -- they
involve patterns over the transitive closure of containment links in pre-condition graphs.

Post-conditions also use the same pattern language as the
apply graphs of DSLTrans transformation rules, with the additional
possibility of expressing indirect links in post-conditions. Traceability links can
also be used in properties to impose traceability relations between pre-condition
and post-condition elements.

Note that \cref{def:trans_prop_appendix} includes a condition stating a surjective typed graph homomorphism needs to exist between the match part of at least one of the transformation's path condition, and the pre-condition of the property of interest. This condition makes sure that the property's pre-condition can be found at least in one execution of the transformation abstracted (the mathematical argument for this fact is given in the proof of \cref{prop:proof_validity}). This condition makes the checking the validity of a property in the transformation meaningful. If this condition would not be true then it could be that the input pattern required by the property would never be fully matched during transformation execution, making such a property not relevant\footnote{In~\cite{DBLP:conf/sle/BarrocaLAFS10} we have referred to these properties \emph{non-provable}. In the work presented here we explicitely disallow the construction of this class of properties.} for the transformation at hand.

\begin{definition}{Satisfaction of a Property by an Execution of a Transformation\\}
\label{def:sat_prop_ex_appendix}
\CatchFileBetweenTags{\satpropexe}{text/definitions}{satpropexe}{\satpropexe}
\end{definition}

\Cref{def:sat_prop_ex_appendix} states that, every time a graph that is isomorphic to the property's pre-condition is found in (the containment transitive closure of) the input model of the transformation's execution, a graph that is isomorphic to the complete property needs to be found in (the containment transitive closure of) the transformation execution. Note that the last part of the proposition in \cref{def:sat_prop_ex_appendix} ensures that the graph that is isomorphic to the property's pre-condition and the graph that is isomorphic to the complete property overlap on their pre-condition parts.

% resulting from the input pattern of $ex$ identified as pre-condition by $f$. Generally speaking, whenever the property's pre-condition is isomorphically found as a pattern of the input model of a execution, the property's post-condition must also be isomorphically found in the output generated from that input pattern. Note that, because a property may contain traceability links between the pre- and post-condition elements, it is necessary to check that those traceability links are also present in the execution. Because of this, the $g$ homomorphism in definition~\ref{def:sat_prop_ex} that checks the property's post-condition is such that it considers the same match pattern identified by the $f$ homomorphism when the pre-condition is checked.

\begin{definition}{Satisfaction of a Property by a Path Condition\\}
\label{def:sat_prop_pc_appendix}
\CatchFileBetweenTags{\satproppc}{text/definitions}{satproppc}{\satproppc}
\end{definition}

The principle behind the satisfaction relation in \cref{def:sat_prop_pc_appendix} is the same as the one behind the satisfaction relation between a property and an execution of a transformation in \cref{def:sat_prop_ex_appendix}: whenever the property's pre-condition is found in the path condition then so is the complete property. Also, those two graphs found in the path condition share the property's pre-condition part. This last condition enforces that the pre- and post-conditions of the property are correctly linked by traceability link in the path condition.

\begin{proposition}{(Validity) The result of proving a property on a set of path conditions generated for a transformation or an all executions of that transformation is the same.\\}
\label{prop:proof_validity_appendix}
Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a transformation and $p \in \textsc{Property}(tr)$ be a
property of $tr$. This given, we have that transformation $tr$ satisfies property $p$ if and only if:
\begin{equation}
\label{eq:prop_proof_transf_appendix}
\bigwedge_{pc\in \textsc{Pathcond}(tr)} pc\vdash p \hspace{.3cm} \Longleftrightarrow \hspace{.3cm} \bigwedge_{ex\in \textsc{Exec}(tr)}ex\models p
\end{equation}
\end{proposition}
\begin{pf}
In order to prove the proposition in \cref{eq:prop_proof_transf_appendix} we will start by demonstrating that,
if property $p$ holds on a path condition $pc$ generated for $tr$,
then $p$ will necessarily hold on any execution $ex$ of $tr$ that is abstracted by $pc$. On the other hand, if $p$ does not hold on $pc$ then it will not hold for at least of one execution $ex$ of $tr$ abstracted by $pc$. This lemma can be stated as follows:
\begin{equation}
\label{eq:prop_proof_pc_appendix}
pc\vdash p \;\Longleftrightarrow \;\forall ex\in\{ex\in\textsc{Exec}(tr)\;|\;ex\Vvdash pc\}\;.\; ex\models p
\end{equation}
We thus need to demonstrate both directions of the equivalence in \cref{eq:prop_proof_pc_appendix}. On the one hand we need to prove of the left-to-right direction of the equivalence:
\begin{multline}
\label{eq:prop_proof_left_right_appendix}
pc\vdash p \;\Longrightarrow \;\forall ex\in\{ex\in\textsc{Exec}(tr)\;|\;ex\Vvdash pc\}\;.\; ex\models p
\end{multline}
Proposition~\ref{eq:prop_proof_left_right_appendix} is shown to be true in \cref{lemma:validity1_appendix}.
We then need to show the right-to-left direction of the equivalence:
\begin{multline}
\label{eq:prop_proof_right_left_appendix}
\forall ex\in\{ex\in\textsc{Exec}(tr)\;|\;ex\Vvdash pc\}\;.\; ex\models p \;\Longrightarrow \;pc\vdash p 
\end{multline}
Proposition~\ref{eq:prop_proof_right_left_appendix} is shown to be true in \cref{lemma:validity2_appendix}.
Once propositions~\ref{eq:prop_proof_left_right_appendix} and~\ref{eq:prop_proof_right_left_appendix} are proved, we know that all path conditions on which a property holds represent executions on which the property also holds. Thus, if the property holds on all path conditions then it necessarily holds on all executions. On the other hand, if a property does not hold on one path condition, making it such that the conjunction on the left side of the equivalence in \cref{eq:prop_proof_transf_appendix} is false, then according to \cref{eq:prop_proof_pc_appendix} an execution for which it also does not hold exists. This makes it such that the conjunction on the right side of the equivalence in \cref{eq:prop_proof_pc_appendix} is also false.
\end{pf}

@@

\begin{proposition}{(Completeness) Properties of a transformation can be shown to either hold for all transformation executions, or not hold for at least one transformation execution.}
\label{prop:proof_completeness_appendix}
\end{proposition}
\begin{proof}
This result follows from two previous results: \cref{prop:pc_completeness_appendix}, that tells us that every transformation execution is abstracted by one path condition; and \cref{prop:proof_validity_appendix} that shows us that every path condition is taken into consideration during property proof. Note that \cref{lem:uniqueness_appendix} guarantees consistency of our results, in the sense that the uniqueness of one path condition per transformation execution guarantees that a property cannot be proven to be both \emph{true} and \emph{false} for two path conditions representing the same transformation execution.
\end{proof}

\clearpage

>>>>

\subsection{Validity and Completeness}
\label{subsec:prop_proving_valid_complete}

As for the path condition building algorithm, \emph{validity} and
\emph{completeness} need to be examined regarding our property verification
algorithm. In this context \emph{validity} means that if a property is satisfied by all path conditions generated for a transformation $tr$, then that property is satisfied by all executions of that transformation. On the other hand, if the property is not satisfied by at least one path condition, then it will not be satisfied by at least one transformation execution. In other words, we wish to show that no false positive or false negative proof results are induced by the abstraction relation.

% The proof of of these statements necessarily relies on the abstraction relation formally presented in\cref{def:instance_pc_ex}.

On the other hand, \emph{completeness} means that we are sure that all properties that can be expressed about a transformation can be shown to hold or not hold in all transformation executions.

As with the proofs for the validity and completeness of the abstraction relation, we present only proof sketches in this section in the interest of readability. Full proofs are shown in \cref{sec:val_complete_prop_verif} as \cref{prop:proof_validity_appendix} and \cref{prop:proof_completeness_appendix}.

% \subsubsection{Validity}
% 
% In this section we prove the validity of our property proof algorithm by considering what it
% formally means when a property is proved on a path condition. We argue that if
% the property holds or not on the path condition, then it will accordingly hold
% or not on for any transformation executions that the path condition represents.

% As well, we also prove that if a property holds or not on a specific transformation
% execution, it will hold or not on the path condition that represents that
% transformation execution.
\begin{proposition}{(Validity) The result of proving a property on a set of path conditions generated for a transformation or an all executions of that transformation is the same.\\}
\label{prop:proof_validity_appendix}
Let $tr\in \textsc{Transf}^{sr}_{tg}$ be a transformation and $p \in \textsc{Property}(tr)$ be a
property of $tr$. This given, we have that transformation $tr$ satisfies property $p$ if and only if:
\begin{equation}
\label{eq:prop_proof_transf_appendix}
\bigwedge_{pc\in \textsc{Pathcond}(tr)} pc\vdash p \hspace{.3cm} \Longleftrightarrow \hspace{.3cm} \bigwedge_{ex\in \textsc{Exec}(tr)}ex\models p
\end{equation}
\end{proposition}
\begin{pf}
In order to prove the proposition in \cref{eq:prop_proof_transf_appendix} we will start by demonstrating that,
if property $p$ holds on a path condition $pc$ generated for $tr$,
then $p$ will necessarily hold on any execution $ex$ of $tr$ that is abstracted by $pc$. On the other hand, if $p$ does not hold on $pc$ then it will not hold for at least of one execution $ex$ of $tr$ abstracted by $pc$. This lemma can be stated as follows:
\begin{equation}
\label{eq:prop_proof_pc_appendix}
pc\vdash p \;\Longleftrightarrow \;\forall ex\in\{ex\in\textsc{Exec}(tr)\;|\;ex\Vvdash pc\}\;.\; ex\models p
\end{equation}
We thus need to demonstrate both directions of the equivalence in \cref{eq:prop_proof_pc_appendix}. On the one hand we need to prove of the left-to-right direction of the equivalence:
\begin{multline}
\label{eq:prop_proof_left_right_appendix}
pc\vdash p \;\Longrightarrow \;\forall ex\in\{ex\in\textsc{Exec}(tr)\;|\;ex\Vvdash pc\}\;.\; ex\models p
\end{multline}
Proposition~\ref{eq:prop_proof_left_right_appendix} is shown to be true in \cref{lemma:validity1_appendix}.
We then need to show the right-to-left direction of the equivalence:
\begin{multline}
\label{eq:prop_proof_right_left_appendix}
\forall ex\in\{ex\in\textsc{Exec}(tr)\;|\;ex\Vvdash pc\}\;.\; ex\models p \;\Longrightarrow \;pc\vdash p 
\end{multline}
Proposition~\ref{eq:prop_proof_right_left_appendix} is shown to be true in \cref{lemma:validity2_appendix}.
Once propositions~\ref{eq:prop_proof_left_right_appendix} and~\ref{eq:prop_proof_right_left_appendix} are proved, we know that all path conditions on which a property holds represent executions on which the property also holds. Thus, if the property holds on all path conditions then it necessarily holds on all executions. On the other hand, if a property does not hold on one path condition, making it such that the conjunction on the left side of the equivalence in \cref{eq:prop_proof_transf_appendix} is false, then according to \cref{eq:prop_proof_pc_appendix} an execution for which it also does not hold exists. This makes it such that the conjunction on the right side of the equivalence in \cref{eq:prop_proof_pc_appendix} is also false.
\end{pf}

@@

\begin{proposition}{(Completeness) Properties of a transformation can be shown to either hold for all transformation executions, or not hold for at least one transformation execution.}
\label{prop:proof_completeness}
\end{proposition}
\begin{pf}
This result follows from two previous results: \cref{prop:pc_completeness}, that tells us that every transformation execution is abstracted by one path condition; and \cref{prop:proof_validity} that shows us that every path condition is taken into consideration during property proof. Note that \cref{lem:uniqueness} guarantees consistency of our results, in the sense that the uniqueness of one path condition per transformation execution guarantees that a property cannot be proven to be both \emph{true} and \emph{false} for two path conditions representing the same transformation execution.
\end{pf}
